A Case Study in Preserving a High Energy Physics Application

### 2014-02-04 paper.tex
modify version-evolution pic
add related work part

\title{A Case Study in Preserving a High Energy Physics Application}
\author{
Haiyan Meng, Matthias Wolf, Peter Ivie, Anna Woodard, Michael Hildreth, and Douglas Thain\\
\affaddr{Department of Physics and Department of Computer Science and Engineering}\\
\affaddr{University of Notre Dame}\\
\affaddr{\{hmeng|mwolf3|awoodard|pivie|mhildreth|dthain\}@nd.edu}
}
\date{15 January 2014}
\maketitle

@@

\begin{figure}[t]
\centering
\includegraphics[width=.8\columnwidth]{data-code-size.eps}
\caption{Inputs to Tau Analysis}
\label{fig:data-code-size}
\end{figure}

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{|l|l|r|r|r|}
        \hline
        \bf Name & \bf Location & \bf Total & \bf Named & \bf Used \\ 
        \hline
        SBEAN data    & hdfs & 33TB & 43.3GB & 20GB \\ \hline
        CMSSW binaries     & local & 1.16TB & 449MB & 5.2MB\\ \hline
        Tau source       & git & 51MB & 51MB & 51MB \\ \hline
        PyYAML binaries    & http & 52MB & 51MB & 51MB \\ \hline
        .h file       & http & 41KB & 41KB & 41KB \\ \hline
        Misc commands & hpc fs & 206TB & ??? & 119MB \\ \hline
        Linux commands & local fs & ??? & ??? & 88.4MB \\ \hline     
        Total          &          & ??? & ??? & ??? \\ \hline
    \end{tabular}
    \caption{Data and Code Used by Tau Analysis}
    \label{table:size-original-real}
\end{table}

@@

The application which is the study of this paper is called \emph{TauAnalysis}
It searches for cases where the the Higgs boson decays to two tau leptons.
The leptons are not observed directly, but by the particle showers
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

Figure~\ref{fig:data-code-size} shows that both the code and data
that form \emph{TauAnalysis} are drawn from large repositories through
multiple steps of reduction.

@@

In the second step, which is the beginning of the actual tau analysis,
the data are reduced to variables relevant to the tau analysis, while
only events matching basic quality criteria are kept.  This results in
a dataset of 43.3$\,$GB.  Again, the Notre Dame CMS groups cluster
resources are used to perform this reduction and selection,
running highly customized software,
built on CMSSW and the BEAN framework,
but with output code written and maintained by a few people only.
Again, the code is stored in a git repository.

The final data analysis, investigated below, can be run as a single
process, and contains a stringent event selection to keep only high
quality candidate events for the underlying physical process (using
122$\,$MiB of space).  Quantities from the selected events can be
both plotted and used in multivariate analysis to determine the level
of expected signal in real data.
This package is written using the CMSSW build framework,
but only utilizes code from ROOT,
a particle physics toolkit underlying CMSSW,
and a few external python dependencies for convenience.
The latter have to be manually fetched and installed,
while the analysis program is built by CMSSW after being checked out of git.

{\bf Code Sources.} Like many scientific codes, the central algorithm
of \emph{TauAnalysis} is expressed in a relatively small amount of
custom code developed by the primary author.  But, the code cannot
run at all without making use of an enormous collection of software
dependencies.  Some of these dependencies are standard to operating
systems worldwide, some are standardized across the entire high-energy
physics field, some are particular to small collaborative groups,
and a few are very specific to a single researcher.

@@

\emph{TauAnalysis} was provided to us in the form of an
email which described, in prose, how to obtain the source,
build the program, and run it correctly on one specific
machine at Notre Dame, with no particularly guarantee that
it will run anywhere else in the world.
Although this starting point may seem extreme, it is
perfectly natural for collaborators to share configurations
with each other in this form, and to rely on the presence
of a working environment with appropriate dependencies already
installed.

@@

@@

\subsubsection{Size of Data and Software Dependencies}
The size of data and software dependencies is astonishing.
The size of BEAN is 11.6TB, which is too large to be analyzed directly by the tau analysis program. After one preprocessing which only collects the events matching basic quality criteria, the data size is still very large, 43.3GB.

@@

\subsubsection{Complexity of the Program}
Complex data and software dependencies, complex environment building process, the large size of data and software involved in the tau analysis, together with the access authority problem, determine the complexity of the program. 
In order to obtain the experimental data from the Grid, the author needs to install and config grid-control software. 
The access authority of the Grid is also necessary for the data acquisition.
The software acquisition from CMSSW, Git and HTTP and environment building process is time-consuming. The execution time of actual data analysis is about 30 minutes. However, the software acquisition and environment building time, excluding data acquisition time, is about 30 minutes, which is very close to the execution time of actual data analysis. If the acquisition of experimental data is considered, actual data analysis only occupies one small percentage of the whole time consumption.
The size of data and software dependencies further complicate the program.

@@

Even if the orginal data size referred by the program is astonishing, the size
of really used data and software is greatly smaller.
Figure~\ref{fig:data-code-size} illustrates the procedures of the tau analysis.
After SBEAN, the data size is 43.3GB, however, the atual size of really used
data in the program is 20GB.  The same trend occurs to the size of software and
other dependencies. The size of CMSSW repository is 1.16TB, the size of
packages checked out from CMSSW is 449MB, and the size of actually used CMSSW
files is only 5.2MB. 
The total size of PanFS mounted as /pscratch is 206TB, however, the really used data from /pscratch is only 119MB.
We compared the original size and actually used size of data and software in Table~\ref{table:size-original-real}.

@@

\begin{table}
    \centering
    \begin{tabular}{|l|}
        \hline
        Version 1: Email\\ \hline
        \\
        1. Create CMS release,\\
            \hspace{9pt} e.g. cmsrel CMSSW\_5\_3... \\
        2. Install BEAN packages: \\
            \hspace{9pt} https://github.com/cms-ttH/...\\
        3. Install grid-control: \\ 
            \hspace{9pt} svn co https://dwekptrack... \\
        4. INstall the TauAnalysis package: \\
           \hspace{9pt} git clone https://github.com/matz-e/... \\
           \hspace{9pt} scram b -j32 \\
        5. Fix grid\_control.cfg and run it. \\
        6. Perform actual analysis \\ 
        \\ \hline
        Version 2: Script\\ \hline
        \\
        set CMSSW\_BASE = (CMSSW\_5\_3...) \\
        module load git \\        
        cmsenv \\ 
        cvs login \\
        cvs co UserCode/Bicocca/... \\
        git clone... \\
        scram b -j32 \\
        wget http://pyyaml.org/... \\
        python setup.py install... \\
        \#the experiment data is from HDFS \\
        roaster data/generic\_ttl.yaml \\ 
        \\ \hline
        Version 3: Fine-Grained Toolkit\\ \hline
        \\
        \#the description of OS version and hardware architecture \\
        \# list of indirect software dependencies \\
        set CMSSW\_BASE = (CMSSW\_5\_3\_11\_patch3) \\ 
        cmsenv \\
        wget http://pyyaml.org/... \\
        python setup.py install... \\
        roaster data/generic\_ttl.yaml \\
        \\ \hline
        Version 4\\ \hline
        \\
        \#the description of OS version and hardware architecture \\
        \#declaration of environment variables\\
        \#data dependencies\\
        \#direct and indirect software dependencies\\
        \#software build and install commands\\
        \#actual analysis program\\ 
        \\ \hline
    \end{tabular}
    \caption{Scripts of each Solution}
    \label{table:scripts}
\end{table}

@@

\subsection{Solution 1: Email}
In order to repeat Matthais's example, the new user consulted the original author by email about the necessary work for the experiment. In response, Matthias introduced the general workflow of his tau analysis through one long email including notes, linux shell commands, web links. The Version 1 of Table~\ref{table:scripts} illstrates his email.

However, this solution to repeat one experiment has three potential drawbacks.
Firstly, the experience of repeating one tau analysis through emails is
chaotic. You need to constantly jump around multiple web links. There are overlap
between the content of the email and the content of web links, which needs the
new user to merge them. Multiple communication through emails is necessary to
ensure the successful reproduction of the whole analysis. For example, the original
email refers to one environment variable called CMSSW\_base without clear
declaration, the new user needs to send one email to the original writer to
obtain its accurate value. Secondly, the necessary procedures to repeat the
experiment, including software acquisition from different sources, is complex
for the new user. In Matthias's example, the sources of software includes
CMSSW, Git, HTTP. The access of CMSSW requires the new user to be an authorized
user of CMSSW. What's more, some parts of the workflow are unrepeatable. The
third step of this experiment requires the new user to own the access authority
of the Grid.

@@

\section{Related Work }

@@

this is an example of citing \cite{Laboratories79make}.

>>>>

\title{A Case Study in Preserving a High Energy Physics Application}
\author{
Haiyan Meng, Matthias Wolf, Peter Ivie, Anna Woodard, Michael Hildreth, and Douglas Thain\\
\affaddr{Department of Physics and Department of Computer Science and Engineering}\\
\affaddr{University of Notre Dame}\\
\affaddr{\{hmeng|mwolf3|pivie|awoodard|mhildreth|dthain\}@nd.edu}
}
\date{15 January 2014}
\maketitle

@@

\begin{figure}[t]
\centering
\includegraphics[width=.8\columnwidth]{data-code-size.eps}
\caption{Inputs to Tau Roast}
\label{fig:data-code-size}
\end{figure}

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{|l|l|r|r|r|}
        \hline
        \bf Name & \bf Location & \bf Total & \bf Named & \bf Used \\ 
        \hline
        Ntuples data    & hdfs & 33TB & 43.3GB & 20GB \\ \hline
        CMSSW binaries     & local & 1.16TB & 449MB & 5.2MB\\ \hline
        Tau source       & git & 51MB & 51MB & 51MB \\ \hline
        PyYAML binaries    & http & 52MB & 51MB & 51MB \\ \hline
        .h file       & http & 41KB & 41KB & 41KB \\ \hline
        Misc commands & hpc fs & 206TB & ??? & 119MB \\ \hline
        Linux commands & local fs & ??? & ??? & 88.4MB \\ \hline     
        Total          &          & ??? & ??? & ??? \\ \hline
    \end{tabular}
    \caption{Data and Code Used by Tau Roast}
    \label{table:size-original-real}
\end{table}

@@

The application which is the study of this paper is called \emph{TauRoast}
It searches for cases where the the Higgs boson decays to two tau leptons.
The leptons are not observed directly, but by the particle showers
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

Figure~\ref{fig:data-code-size} shows that both the code and data
that form \emph{TauRoast} are drawn from large repositories through
multiple steps of reduction.

@@

In the second step, which is the beginning of the actual tau analysis,
the data are reduced to variables relevant to the tau roast procedure, while
only events matching basic quality criteria are kept.  This results in
a dataset of 43.3$\,$GB.  Again, the Notre Dame CMS groups cluster
resources are used to perform this reduction and selection,
running highly customized software,
built on CMSSW and the BEAN framework,
but with output code written and maintained by a few people only.
Again, the code is stored in a git repository.

The final data analysis, investigated below, can be run as a single
process, and contains a stringent event selection to keep only high
quality candidate events for the underlying physical process (using
about 20$\,$MiB of space).  Quantities from the selected events can be
both plotted and used in multivariate analysis to determine the level
of expected signal in real data.
This package is written using the CMSSW build framework,
but only utilizes code from ROOT,
a particle physics toolkit underlying CMSSW,
and a few external python dependencies for convenience.
The latter have to be manually fetched and installed,
while the analysis program is built by CMSSW after being checked out of git.

{\bf Code Sources.} Like many scientific codes, the central algorithm
of \emph{TauRoast} is expressed in a relatively small amount of
custom code developed by the primary author.  But, the code cannot
run at all without making use of an enormous collection of software
dependencies.  Some of these dependencies are standard to operating
systems worldwide, some are standardized across the entire high-energy
physics field, some are particular to small collaborative groups,
and a few are very specific to a single researcher.

@@

\emph{TauRoast} was provided to us in the form of an
email which described, in prose, how to obtain the source,
build the program, and run it correctly on one specific
machine at Notre Dame, with no particularly guarantee that
it will run anywhere else in the world.
Although this starting point may seem extreme, it is
perfectly natural for collaborators to share configurations
with each other in this form, and to rely on the presence
of a working environment with appropriate dependencies already
installed.

@@

\begin{figure}[t]
\centering
\includegraphics[width=.8\columnwidth]{dependencies.eps}
\caption{Dependencies}
\label{fig:dependencies}
\end{figure}

@@

\subsubsection{Size of Data and Software Dependencies}
The size of data and software dependencies is astonishing.
The size of BEAN is 11.6TB, which is too large to be analyzed directly by the tau roast program. After one preprocessing which only collects the events matching basic quality criteria, the data size is still very large, 43.3GB.

@@

\subsubsection{Complexity of the Program}
Complex data and software dependencies, complex environment building process, the large size of data and software involved in the tau roast program, together with the access authority problem, determine the complexity of the program. 
In order to obtain the experimental data from the Grid, the author needs to install and config grid-control software. 
The access authority of the Grid is also necessary for the data acquisition.
The software acquisition from CMSSW, Git and HTTP and environment building process is time-consuming. The execution time of actual data analysis is about 30 minutes. However, the software acquisition and environment building time, excluding data acquisition time, is about 30 minutes, which is very close to the execution time of actual data analysis. If the acquisition of experimental data is considered, actual data analysis only occupies one small percentage of the whole time consumption.
The size of data and software dependencies further complicate the program.

@@

Even if the orginal data size referred by the program is astonishing, the size
of really used data and software is greatly smaller.
Figure~\ref{fig:data-code-size} illustrates the inputs of the tau roast program.
After the TauAnalysis procedure, the data size of Ntuples is 43.3GB, however, the atual size of really used
data in the program is 20GB.  The same trend occurs to the size of software and
other dependencies. The size of CMSSW repository is 1.16TB, the size of
packages checked out from CMSSW is 449MB, and the size of actually used CMSSW
files is only 5.2MB. 
The total size of PanFS mounted as /pscratch is 206TB, however, the really used data from /pscratch is only 119MB.
We compared the original size and actually used size of data and software in Table~\ref{table:size-original-real}.

@@

\begin{table}
    \centering
    \begin{tabular}{|l|}
        \hline
        Version 1: Email\\ \hline
        \\
        1. Create CMS release,\\
            \hspace{9pt} e.g. cmsrel CMSSW\_5\_3... \\
        2. Install BEAN packages: \\
            \hspace{9pt} https://github.com/cms-ttH/...\\
        3. Install grid-control: \\ 
            \hspace{9pt} svn co https://dwekptrack... \\
        4. INstall the TauAnalysis package: \\
           \hspace{9pt} git clone https://github.com/matz-e/... \\
           \hspace{9pt} scram b -j32 \\
        5. Fix grid\_control.cfg and run it. \\
        6. Perform actual tau roast program \\ 
        \\ \hline
        Version 2: Script\\ \hline
        \\
        set CMSSW\_BASE = (CMSSW\_5\_3...) \\
        module load git \\        
        cmsenv \\ 
        cvs login \\
        cvs co UserCode/Bicocca/... \\
        git clone... \\
        scram b -j32 \\
        wget http://pyyaml.org/... \\
        python setup.py install... \\
        \#the experiment data is from HDFS \\
        roaster data/generic\_ttl.yaml \\ 
        \\ \hline
        Version 3: Fine-Grained Toolkit\\ \hline
        \\
        \#the description of OS version and hardware architecture \\
        \# list of indirect software dependencies \\
        set CMSSW\_BASE = (CMSSW\_5\_3\_11\_patch3) \\ 
        cmsenv \\
        wget http://pyyaml.org/... \\
        python setup.py install... \\
        roaster data/generic\_ttl.yaml \\
        \\ \hline
        Version 4\\ \hline
        \\
        \#the description of OS version and hardware architecture \\
        \#declaration of environment variables\\
        \#data dependencies\\
        \#direct and indirect software dependencies\\
        \#software build and install commands\\
        \#actual analysis program\\ 
        \\ \hline
    \end{tabular}
    \caption{Scripts of each Solution}
    \label{table:scripts}
\end{table}

@@

\subsection{Solution 1: Email}

In order to repeat Matthais's example, the new user consulted the original
author by email about the necessary work for the experiment. In response,
Matthias introduced the general workflow of his tau roast program through one long
email including notes, linux shell commands, web links. The Version 1 of
Table~\ref{table:scripts} illstrates his email.

However, this solution to repeat one experiment has three potential drawbacks.
Firstly, the experience of repeating one tau roast program through emails is
chaotic. You need to constantly jump around multiple web links. There are overlap
between the content of the email and the content of web links, which needs the
new user to merge them. Multiple communication through emails is necessary to
ensure the successful reproduction of the whole analysis. For example, the original
email refers to one environment variable called CMSSW\_base without clear
declaration, the new user needs to send one email to the original writer to
obtain its accurate value. Secondly, the necessary procedures to repeat the
experiment, including software acquisition from different sources, is complex
for the new user. In Matthias's example, the sources of software includes
CMSSW, Git, HTTP. The access of CMSSW requires the new user to be an authorized
user of CMSSW. What's more, some parts of the workflow are unrepeatable. The
third step of this experiment requires the new user to own the access authority
of the Grid.

@@

\section{Related Work }
\subsection{Software Preservation Approaches}

Generally, there are three approaches to preserve software environment:
technical/hardware preservation, migration and emulation.  Hardware
preservation preserves the original software and its original operating
environment. \cite{cifuentes1996binary} presented techniques used to migrate
legacy software running on register-based machines to modern RISC machines. \cite{mancl2001refactoring}
utilized one method called refactoring to make software migration become
easier. However, migration often involves the re-compiling and re-configuring
the source code to accustom a new hardware platform and software environment.
Emulation recreate the original software and hardware environment by
programming future platforms and OSs. One common solution to implement this is
virtual machine. According to the usage and emulation degree of the real
machine, virtual machine can be divided into system virtual machine and process
virtual machine. \cite{goldberg1974survey}\cite{smith2005architecture} introduced the working principle, design principle and
performance evaluation of system virtual machine. \cite{barham2003xen}\cite{kivity2007kvm}\cite{rosenblum1999vmware} illustrated the
functionality of  system VM to support different guest operating systems. \cite{esquembre2004easy} illustrated how JVM, one process virtual machine, can expedite the creation of
scientific simulations in Java. \cite{matthews2009towards}\cite{phelps2005no}\cite{hong2010software} discussed the pros and cons of each
approach.

\subsection{Software Environment vs Computing Environment}

\cite{matthews2009towards}\cite{phelps2005no}\cite{hong2010software} treated the preservation of computing environment and software environment as one entirety. However, frequently changing experiment software makes the maintenance of the preserved experimental environment very complex.

\cite{buncic2010cernvm} treated them as two different categories. The preservation of computing environment is implemented with CernVM, and the preservation of software environment is based on a CernVM filesystem(CVMFS) specifically designed for efficient software distribution.

\subsection{Software Preservation Format}

\cite{zabolitzky2002preserving}\cite{castagne2013consider} emphasized the importance of preserving software in source code format.

\cite{buncic2010cernvm} published pre-built and configured experiment software releases to avoid repeating the time-consuming software building procedure.

\subsection{facilitate the reproduction of scientific experiments}

Attempts from different perspectives to facilitate the reproduction of scientific experiments utilizing preserved software library has been made. 
\cite{compostella2010cdf} discussed software distribution mechanism over the Grid using Parrot. 
\cite{blomer2011cernvm} proposed CVMFS to facilitate the software distribution over the network to CernVM. 
\cite{rice1996scientific} made the reproduction process easier through the integration of user interface, scientific software libraries, knowledge base into problem-solving environment.
\cite{kohn2001divorcing} tried to enable the creation and distribution of language-independent software library by addressing language interoperability.

\subsection{figure out the absolutely relevant data and software }

Current mechanisms of preserving scientific experiments assume that all the data and software mentioned in the experiments are necessary for the reproduction of the experiments. However, this is not always right. In some cases, the original author may leave additional code referring to irrelative data and software in the experiment programs. One mechanism, which can figure out the absolutely relevant data and software of one experiment, is important for both the preservation and reproduction of scientific experiments.

\subsection{the importance of case study}
\cite{matthews2008significant} introduced one conceptual framework for software preservation from several case studies of software preservation
\cite{matthews2010framework} designed a tool to capture software preservation properties within a software environment through a series of case studies conducted to evaluate the software preservation framework.
\cite{johnston2014workflow} proposed one overall data curation workflow for 3-5 case studies of preserving research data.
To figure out how to preserve HEP applications, this paper studies one case of preserving one representative HEP application.

@@
